---
title: "COMPSCIX 415.2 Homework 2"
author: "Elena Tolpygo Cranley"
date: "2/06/2018"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

3.2.4 Exercises

- Run ggplot(data = mpg). What do you see?

This creates an empty graph. To show information on the graph, layers/options must be specified.

- How many rows are in mpg? How many columns?
```{r}
glimpse(mpg)
```
234 rows, 11 columns

- What does the drv variable describe? Read the help for ?mpg to find out.

The drv variable describes the type of vehicle drivetrain: f = front-wheel drive, r = rear wheel drive, 4 = 4wd

- Make a scatterplot of hwy vs cyl.
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```


- What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

The plot is not useful because both class and drv are categorical rather than numeric data types.


3.3.1 Exercises

- What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Points are not blue because the color parameter should be outside the inner parentheses.


- Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

Continuous variables are: cty, hwy, displ. All other variables are categorical. 


- Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color=displ))
```

For continuous variables, color is a gradation of shading, size is a range of sizes, and shape cannot be mapped to.

- What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color=hwy, size=hwy))
```

If the same variable is mapped to multiple aesthetics, the same range of values is shown in multiple ways (for example, y-value, size, shading)


- What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

The stroke aesthetic defines border width of shapes. It works with points, triangles, and squares.


- What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

Those values that match the statement will be of one color, and ones that don't, of the other.


3.5.1 Exercises

- Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Faceting allows to see trends within each particular data subset easier; however, it may make it harder to see the overall data trends. For larger data sets, it may be more important to see the facet trends in addition to the overall dataset trends.


- Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?

nrow controls the number of rows in the facets grid, and ncol, the number of columns. Other options controlling the individual panel layout are: scales (whether fixed, or free, or free in one dimension), shrink (whether to shrink scales to fit output of statistics rather than raw data), labeller (how to display the labels), switch (whether to display labels on the top and right of the plot), strip.position (position of the labels)


3.6.1 Exercises

- What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

line chart: geom_line()

boxplot: geom_boxplot()

histogram: geom_histogram()

area chart: geom_area()


- Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

the plot should draw points for each displ/hwy observation, with point color representing the drv. then, a smooth line for each of the drv groupings should be fitted to the same data.


- What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

show.legend = FALSE should remove the legend, but doesn't seem to.


- What does the se argument to geom_smooth() do?

the se argument controls whether the confidence interval (as grey area) should be displayed around the smooth line.


- Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

The graphs should look the same, because they have the same arguments. in the first case, the arguments are specified to the whole ggplot, and in the second, to each individual geom layer.


3.7.1 Exercises


- What does geom_col() do? How is it different to geom_bar()?

geom_col() makes heights of the bars represent values in the data, whereas geom_bar() makes heights of the bar proportional to the number of cases in each group.

~~~~~~~~~~~~~~~~~~~~~~~

Look at the data graphics at the following link: What is a Data Scientist. Please briefly critique the
designer’s choices. What works? What doesn’t work? What would you have done differently?

The designer uses pie charts, which is a questionable choice, but works okay here because the slices are well-labelled. 
There are two sections that do not work:

- Sources of new data science talent

- Data science vs. business intelligence educational background.

The reason these sections don't work well is because the sizes of shapes are disproportionate to their relative values (for example, the 24% square is about 4 times the size of 12% square, and 37% circle is something like 10 times bigger than the 10% circle)

Additionally, the Who does a data scientist work with section is shown using squares, but they are all of the same size, despite percentage values being different. This section could be shown as a simple list or as a bar chart with bar sizes reflecting the percentage values.