---
title: "COMPSCIX 415.2 Homework 1"
author: "Elena Tolpygo Cranley"
date: "1/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/etolpygo/compscix-415-2-assignments](https://github.com/etolpygo/compscix-415-2-assignments)

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

```{r}
data("WorldCities")
```

```{r}
glimpse(WorldCities)
```

QUESTION 1: How many observations and variables are there in this data set? What are some of the variable names? 

There are 23,018 observations of 10 variables. Some of the variables are: code, name, country, latitude, population, region, date.

```{r}
WorldCities <- head(WorldCities, 200)
country_col <- WorldCities$country
unique(country_col)
```

QUESTION 2: There is a column called region in the data set. Can you extract this and show only the unique values?

```{r}
regions_col <- WorldCities$region
unique(regions_col)
```

QUESTION 3: Can you extract and show the unique entries from the country column in WorldCities using one line of code and two %>% operators?

```{r}
WorldCities %>% select(country) %>% unique()
```

```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
```

QUESTION 4: Make one more histogram of the population and add a new x-label, a new ylabel, a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().

```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + xlab('City Population') + ylab('Number of Cities') + ggtitle('Histogram of Populations of World Cities') + theme_bw()
```

